### 2024-11-18 Article/Background_sections/Adhesivity.tex
minima

\begin{definition}(Van Kampen property)
    Let $\mathcal A$ be a subclass of $\Hom(\cat C)$, and consider the diagram below:
    % \[
    % \begin{tikzcd}
    %     A \arrow[r, "f"] \arrow[d, "m" swap] & B \arrow[d, "n"] \\
    %     C \arrow[r, "g" swap]                & D
    % \end{tikzcd}
    % \]
    \[
    \begin{tikzcd}[row sep = scriptsize]
        A' \arrow[ddd, "m'"swap] \arrow[rrr, "f'"] \arrow[dr, "a"swap] & & & B \arrow[ddd, "n'"] \arrow[dl, "b"] \\
        & A \arrow[r, "f"] \arrow[d, "m" swap] & B \arrow[d, "n"]   & \\
        & C \arrow[r, "g" swap]                & D                  & \\
        C' \arrow[rrr, "g'"swap] \arrow[ur, "c"] & & & D' \arrow[ul, "d" swap] 
    \end{tikzcd}
    \]
    \[\begin{tikzcd}[row sep=25, column sep=25]
	& {A'} && {B'} \\
	{C'} && {D'} \\
	& A && B \\
	C && D
	\arrow["{f'}", from=1-2, to=1-4]
	\arrow["{m'}"', from=1-2, to=2-1]
	\arrow["a"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{n'}"', from=1-4, to=2-3]
	\arrow["b", from=1-4, to=3-4]
	\arrow["{g'}"{pos=0.7}, from=2-1, to=2-3, crossing over]
	\arrow["c"', from=2-1, to=4-1]
	\arrow["f"'{pos=0.3}, from=3-2, to=3-4]
	\arrow["d"{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["m", from=3-2, to=4-1]
	\arrow["n", from=3-4, to=4-3]
	\arrow["g"', from=4-1, to=4-3]
    \end{tikzcd}\]
    we say that the inner square is an \emph{$\mathcal A$-Van Kampen} square if:
    \begin{itemize}
        \item it is a pushout;
        \item $a, b, c, d \in \mathcal{A}$;
        \item whenever the top and the left squares are pullbacks then the outer square is a pushout if and only of the right and the bottom squares are pullbacks.
    \end{itemize}
\end{definition}

@@

\begin{proof}
    In order to prove $\mathcal M$-adhesivity of $\cat C$, we have to verify the condition in \Cref{def:adh}.
    \begin{itemize}
        \item Let $C \xrightarrow[]{g} D \xleftarrow[]{m} B$ with $m \in \mathcal M$ be a cospan in $\cat C$. Applying $F$, we obtain $F(C) \xrightarrow[]{F(g)} F(D) \xleftarrow[]{F(m)} B$, with $F(m) \in \mathcal{N}$ by hypothesis. Then, there exists a pullback $(P_F, p_{F(B)}, p_{(F(D))})$ in $\cat D$, which is an $\mathcal N$-pullback (\Cref{def:pullback_pushout}). Since $F$ creates pullbacks, hence lifts them (\Cref{obs:funct_creat_lim_then_lift}), there exist a pullback $(P, p_B, p_D)$ in $\cat C$.
        \item Let $C \xleftarrow{m} A \xrightarrow{f} B$ with $ m \in \mathcal M$ be a cospan in $\cat C$. Analogously to the previous point, applying the functor $F$ we obtain $F(C) \xleftarrow{F(m)} F(A) \xrightarrow{F(f)} F(B)$ with $ F(m) \in \mathcal N$, and there exists a $\mathcal N$-pushout $(q_{F(C)}, q_{F(B)}, F(Q))$ in $\cat D$. Since $F$ reflects pushouts, $(q_C, q_B, Q)$ is a $\mathcal{M}$-pushout in $\cat C$.
        \item the Van Kampen property of $\mathcal M$-pullbacks follows from the closure under pullbacks and pushouts of $\mathcal M$ and from the fact that $F$ reflects pullbacks.
    \end{itemize}

>>>>

\begin{definition}(Van Kampen property)
    Let $\mathcal A$ be a subclass of $\Hom(\cat C)$, and consider the diagram below:
    % \[
    % \begin{tikzcd}
    %     A \arrow[r, "f"] \arrow[d, "m" swap] & B \arrow[d, "n"] \\
    %     C \arrow[r, "g" swap]                & D
    % \end{tikzcd}
    % \]
	\iffalse
    \[
    \begin{tikzcd}[row sep = scriptsize]
        A' \arrow[ddd, "m'"swap] \arrow[rrr, "f'"] \arrow[dr, "a"swap] & & & B \arrow[ddd, "n'"] \arrow[dl, "b"] \\
        & A \arrow[r, "f"] \arrow[d, "m" swap] & B \arrow[d, "n"]   & \\
        & C \arrow[r, "g" swap]                & D                  & \\
        C' \arrow[rrr, "g'"swap] \arrow[ur, "c"] & & & D' \arrow[ul, "d" swap] 
    \end{tikzcd}
    \]
	\fi
    \[\begin{tikzcd}[row sep=25, column sep=25]
	& {A'} && {B'} \\
	{C'} && {D'} \\
	& A && B \\
	C && D
	\arrow["{f'}", from=1-2, to=1-4]
	\arrow["{m'}"', from=1-2, to=2-1]
	\arrow["a"'{pos=0.7}, from=1-2, to=3-2]
	\arrow["{n'}"', from=1-4, to=2-3]
	\arrow["b", from=1-4, to=3-4]
	\arrow["{g'}"{pos=0.7}, from=2-1, to=2-3, crossing over]
	\arrow["c"', from=2-1, to=4-1]
	\arrow["f"'{pos=0.3}, from=3-2, to=3-4]
	\arrow["d"{pos=0.3}, from=2-3, to=4-3, crossing over]
	\arrow["m", from=3-2, to=4-1]
	\arrow["n", from=3-4, to=4-3]
	\arrow["g"', from=4-1, to=4-3]
    \end{tikzcd}\]
    we say that the inner square is an \emph{$\mathcal A$-Van Kampen} square if:
    \begin{itemize}
        \item it is a pushout;
        \item $a, b, c, d \in \mathcal{A}$;
%        \item whenever the top and the left squares are pullbacks then the outer square is a pushout if and only of the right and the bottom squares are pullbacks.
	\item whenever the back and the left squares of the cube are pullbacks, then the top square is a pullback if and only if the front and the right squares are pullbacks.
    \end{itemize}
\end{definition}

@@

\begin{proof}
    In order to prove $\mathcal M$-adhesivity of $\cat C$, we have to verify the condition in \Cref{def:adh}.
    \begin{itemize}
        \item Let $C \xrightarrow[]{g} D \xleftarrow[]{m} B$ with $m \in \mathcal M$ be a cospan in $\cat C$. Applying $F$, we obtain $F(C) \xrightarrow[]{F(g)} F(D) \xleftarrow[]{F(m)} B$, with $F(m) \in \mathcal{N}$ by hypothesis. Then, there exists a pullback $(P_F, p_{F(B)}, p_{F(D)})$ in $\cat D$, which is an $\mathcal N$-pullback (\Cref{def:pullback_pushout}). Since $F$ creates pullbacks, hence lifts them (\Cref{obs:funct_creat_lim_then_lift}), there exist a pullback $(P, p_B, p_D)$ in $\cat C$.
        \item Let $C \xleftarrow{m} A \xrightarrow{f} B$ with $ m \in \mathcal M$ be a cospan in $\cat C$. Analogously to the previous point, applying the functor $F$ we obtain $F(C) \xleftarrow{F(m)} F(A) \xrightarrow{F(f)} F(B)$ with $ F(m) \in \mathcal N$, and there exists a $\mathcal N$-pushout $(q_{F(C)}, q_{F(B)}, F(Q))$ in $\cat D$. Since $F$ reflects pushouts, $(q_C, q_B, Q)$ is a $\mathcal{M}$-pushout in $\cat C$.
        \item the Van Kampen property of $\mathcal M$-pullbacks follows from the closure under pullbacks and pushouts of $\mathcal M$ and from the fact that $F$ reflects pullbacks.
    \end{itemize}
